---
title: "Exploratory data analysis"
author: "Diego Villa"
format: html
editor: visual
---

```{r message=FALSE, warning=FALSE}
library(readr, warn.conflicts = FALSE)
library(skimr)

library(tibble)
library(lubridate, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)
library(slider)
library(scales)
library(tidyr)

library(ggplot2, warn.conflicts = FALSE)
ggplot2::theme_set(theme_bw())
ggplot2::theme_update(panel.grid = element_blank())

library(geofacet)
library(ggsci)
library(GGally)
library(rmcorr)

library(DT)
library(htmltools)
library(stringr)
```

# Data loading 

Create csv. file path. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
processed_data_path <- file.path("../data/processed/")
file_name <- "pro_malaria-monthly-cases-district-loreto"
file <- paste(file_name, ".csv", sep = "")
file_path <- file.path(processed_data_path, file)
```

Define data types for columns.

```{r message=FALSE, warning=FALSE}
# Column names
col_names <- c(
  "district", "year", "month", "falciparum", "vivax", "aet", "prcp", "q",
  "soilm", "tmax", "tmin", "pop2015", "province", "region", "dttm", "time"
)
```

```{r}
# Column types
col_types <- cols(
  district = col_character(), year = col_integer(), month = col_integer(), 
  falciparum = col_integer(), vivax = col_integer(), aet = col_double(), 
  prcp = col_double(), q = col_double(), soilm = col_double(), 
  tmax = col_double(), tmin = col_double(), pop2015 = col_integer(), 
  province = col_character(), region = col_character(), dttm = col_datetime(), 
  time = col_integer()
)
```

Reading csv. file.

```{r message=FALSE, warning=FALSE}
df_malaria_loreto <-
  read_csv(
    file_path,
    col_names = col_names, col_types = col_types, skip = 1,
    locale = locale(encoding = "UTF-8")
  )
```

## Data validation 

TODO: Unique values evaluation.

Check data structure.

```{r message=FALSE, warning=FALSE}
str(df_malaria_loreto)
```

Summarize

```{r summary}
skimr::skim(df_malaria_loreto)
```

Check data structure

```{r data-structure}
str(df_malaria_loreto)
```

```{r}
df_malaria_loreto <- 
  df_malaria_loreto %>% 
  mutate(province = recode(province, PUTUMAYO = "MAYNAS"))
```

Create a data frame with the start and end dates of PAMAFRO intervention.

```{r message=FALSE, warning=FALSE}
df_pamafro_period <- tibble(
  start = make_datetime(year = 2005, month = 10, day = 1L),
  end = make_datetime(year = 2010, month = 09, day = 1L)
)
```

# Exploratory data analysis

TODO: Summarize by districts. `soilm` have the same value for all months in some districts. Do not include this variable in the analysis.

```{r}
df_malaria_loreto %>% 
  dplyr::group_by(year, month) %>% 
  dplyr::summarise(
    vivax      = sum(vivax),
    falciparum = sum(falciparum),
    aet        = mean(aet),
    prcp       = mean(prcp),
    q          = mean(q),
    tmax       = mean(tmax),
    tmin       = mean(tmin)
  ) %>% 
  dplyr::mutate(
    report_date = lubridate::make_datetime(year = year, month = month, day = 1L)
  ) %>% 
  tidyr::pivot_longer(
    cols      = c(vivax, falciparum),
    names_to  = "specie",
    values_to = "count"
  ) %>% 
  dplyr::mutate(
    specie = factor(
      x      = specie, 
      levels = c("vivax", "falciparum"), 
      labels = c("P. vivax", "P. falciparum")
    )
  ) %>% 
  ggplot() +
  geom_line(
    mapping  = aes(x = report_date, y = count, color = specie),
    size     = 1
  ) +
  geom_rect(
    data    = df_pamafro_period,
    mapping = aes(xmin = start, xmax = end), 
    ymin    = -1000, 
    ymax    = 10000,
    alpha   = 0.2
  ) +
  geom_vline(
    xintercept = c(df_pamafro_period$start, df_pamafro_period$end), 
    linetype   = "dashed", 
    alpha      = 0.6
  ) +
  geom_hline(mapping = aes(yintercept = 0), alpha = 0.6, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 8000)) +
  scale_x_datetime(date_labels = "%Y", date_breaks = "1 year") +
  ggsci::scale_color_npg() +
  ggsci::scale_fill_npg() +
  labs(y = "Number of reported cases", x = NULL) +
  theme(legend.position = "top", legend.title = element_blank())
```

```{r message=FALSE}
# df_malaria_loreto %>%
#   dplyr::group_by(year, month) %>%
#   dplyr::summarise(
#     vivax = sum(vivax),
#     falciparum = sum(falciparum),
#     aet = mean(aet),
#     prcp = mean(prcp),
#     q = mean(q),
#     tmax = mean(tmax),
#     tmin = mean(tmin),
#     .groups = "drop"
#   ) %>%
#   dplyr::mutate(
#     report_date = lubridate::make_datetime(year = year, month = month, day = 1L)
#   ) -> df_agg
# 
# df_agg %>% 
#   mutate(
#     aet = rescale(aet, to = c(0, 1)),
#     prcp = rescale(prcp, to = c(0, 1)),
#     q = rescale(q, to = c(0, 1)),
#     tmax = rescale(tmax, to = c(0, 1)),
#     tmin = rescale(tmin, to = c(0, 1))
#   ) -> df_agg_rescale
#   
# 
# df_agg_rescale %>% 
#   tidyr::pivot_longer(
#     cols = c(aet, prcp, q, tmax, tmin),
#     names_to = "variable",
#     values_to = "value"
#   ) %>%
#   dplyr::mutate(
#     variable = factor(
#       x = variable,
#       levels = c("aet", "prcp", "q", "tmax", "tmin"),
#       labels = c(
#         "Actual Evapotranspiration",
#         "Precipitation",
#         "Runoff",
#         "Max Temperature",
#         "Min Temperature"
#       )
#     )
#   ) -> df_agg_rescale_long
# 
# df_agg_rescale_long %>% 
#   ggplot() +
#   geom_line(
#     mapping = aes(x = report_date, y = value, color = variable),
#     size = 1
#   ) +
#   geom_rect(
#     data = df_pamafro_period,
#     mapping = aes(xmin = start, xmax = end),
#     ymin = -1000,
#     ymax = 10000,
#     alpha = 0.2
#   ) +
#   geom_vline(
#     xintercept = c(df_pamafro_period$start, df_pamafro_period$end),
#     linetype = "dashed",
#     alpha = 0.6
#   ) +
#   geom_hline(mapping = aes(yintercept = 0), alpha = 0.6, linetype = "dashed") +
#   scale_y_continuous(limits = c(0, 1)) +
#   scale_x_datetime(date_labels = "%Y", date_breaks = "1 year") +
#   ggsci::scale_color_npg() +
#   ggsci::scale_fill_npg() +
#   labs(y = "Number of reported cases", x = NULL) +
#   theme(legend.position = "top", legend.title = element_blank())
```

```{r message=FALSE}
df_malaria_loreto %>% 
  dplyr::group_by(province, year, month) %>% 
  dplyr::summarise(
    vivax      = sum(vivax),
    falciparum = sum(falciparum),
    pop        = sum(pop2015),
    aet        = mean(aet),
    prcp       = mean(prcp),
    q          = mean(q),
    tmax       = mean(tmax),
    tmin       = mean(tmin)
  ) %>% 
  dplyr::mutate(
    ir_vivax      = vivax / pop,
    ir_falciparum = falciparum / pop
  ) %>% 
  dplyr::group_by(province, year) %>% 
  dplyr::summarise(
    air_vivax      = mean(ir_vivax),
    air_falciparum = mean(ir_falciparum),
    aet            = mean(aet),
    prcp           = mean(prcp),
    q              = mean(q),
    tmax           = mean(tmax),
    tmin           = mean(tmin)
  ) -> df_annual_summary
```

```{r}
df_annual_summary %>% 
  tidyr::pivot_longer(
    cols      = c(air_vivax, air_falciparum),
    names_to  = "specie",
    values_to = "air"
  ) %>% 
  dplyr::mutate(
    specie = factor(
      x      = specie, 
      levels = c("air_vivax", "air_falciparum"), 
      labels = c("P. vivax", "P. falciparum")
    )
  ) -> df_annual_summary_long
```

```{r}
# geofacet::grid_design()
```

```{r}
loreto_grid <- data.frame(
  row  = c(2, 3, 3, 3, 3, 4, 4, 5) - 1,
  col  = c(3, 2, 4, 1, 3, 2, 3, 2),
  code = c("MA", "LO", "PU", "DM", "RC", "AA", "RE", "UC"),
  name = c(
    "MAYNAS", 
    "LORETO", 
    "PUTUMAYO", 
    "DATEM DEL MARAÑON", 
    "MARISCAL RAMON CASTILLA", 
    "ALTO AMAZONAS", 
    "REQUENA", 
    "UCAYALI"
  ),
  stringsAsFactors = FALSE
)
```

## Figure 1

```{r message=FALSE, fig.height=8, fig.width=12, warning=FALSE}
plot_air_province <- 
  ggplot(data = df_annual_summary_long) +
  geofacet::facet_geo(~ province, grid = loreto_grid, label = "name") +
  geom_area(
    mapping  = aes(x = year, y = 10000 * air, color = specie, fill = specie),
    size     = 1,
    alpha    = 0.3,
    position = position_dodge(0)
  ) +
  geom_rect(
    data    = tibble::tibble(start = 2005, end = 2010),
    mapping = aes(xmin = start, xmax = end), 
    ymin    = -100, 
    ymax    = 200,
    alpha   = 0.2
  ) +
  geom_vline(xintercept = c(2005, 2010), linetype = "dashed", alpha = 0.6) +
  scale_y_continuous(limits = c(0, 150)) +
  scale_x_continuous(n.breaks = 5) +
  ggsci::scale_color_npg() +
  ggsci::scale_fill_npg() +
  labs(y = "Annual incidence rate", x = NULL) +
  theme(legend.position = "top", legend.title = element_blank())

plot_air_province
```

```{r message=FALSE}
df_malaria_loreto %>% 
  dplyr::mutate(
    province = recode(province, PUTUMAYO = "MAYNAS")
  ) %>% 
  dplyr::group_by(province, year, month) %>% 
  dplyr::summarise(
    vivax      = sum(vivax),
    falciparum = sum(falciparum),
    pop        = sum(pop2015),
    aet        = mean(aet),
    prcp       = mean(prcp),
    q          = mean(q),
    tmax       = mean(tmax),
    tmin       = mean(tmin),
    .groups = "drop"
  ) %>% 
  dplyr::mutate(
    ir_vivax      = vivax / pop,
    ir_falciparum = falciparum / pop
  ) %>% 
  dplyr::group_by(province, year) %>% 
  dplyr::summarise(
    air_vivax      = mean(ir_vivax),
    air_falciparum = mean(ir_falciparum),
    aet            = mean(aet),
    prcp           = mean(prcp),
    q              = mean(q),
    tmax           = mean(tmax),
    tmin           = mean(tmin),
    .groups = "drop"
  ) -> df_annual_summary_v2
```

```{r}
df_annual_summary_v2 %>% 
  tidyr::pivot_longer(
    cols      = c(air_vivax, air_falciparum),
    names_to  = "specie",
    values_to = "air"
  ) %>% 
  dplyr::mutate(
    specie = factor(
      x      = specie, 
      levels = c("air_vivax", "air_falciparum"), 
      labels = c("P. vivax", "P. falciparum")
    )
  ) -> df_annual_summary_v2_long
```

```{r}
loreto_grid_v2 <- data.frame(
  row  = c(1, 2, 2, 2, 3, 3, 4),
  col  = c(2, 2, 1, 3, 1, 2, 2),
  code = c("MA", "LO", "DM", "RC", "AA", "RE", "UC"),
  name = c(
    "MAYNAS", 
    "LORETO", 
    "DATEM DEL MARAÑON", 
    "MARISCAL RAMON CASTILLA", 
    "ALTO AMAZONAS", 
    "REQUENA", 
    "UCAYALI"
  ),
  stringsAsFactors = FALSE
)
```

```{r}
plot_air_province_v2 <- 
  ggplot(data = df_annual_summary_v2_long) +
  geofacet::facet_geo(~ province, grid = loreto_grid_v2, label = "name") +
  geom_area(
    mapping  = aes(x = year, y = 10000 * air, color = specie, fill = specie),
    size     = 1,
    alpha    = 0.3,
    position = position_dodge(0)
  ) +
  geom_rect(
    data    = tibble::tibble(start = 2005, end = 2010),
    mapping = aes(xmin = start, xmax = end), 
    ymin    = -100, 
    ymax    = 200,
    alpha   = 0.2
  ) +
  geom_vline(xintercept = c(2005, 2010), linetype = "dashed", alpha = 0.6) +
  scale_y_continuous(limits = c(0, 150)) +
  scale_x_continuous(n.breaks = 5) +
  ggsci::scale_color_npg() +
  ggsci::scale_fill_npg() +
  labs(y = "Annual incidence rate", x = NULL) +
  theme(legend.position = "top", legend.title = element_blank(), 
        text = element_text(size = 16))

plot_air_province_v2
```

```{r}
pdf("figs/figure-1.pdf", width = 10, height = 7)
plot_air_province_v2
dev.off()
```



```{r}
plot_vars_ma <- function(prov, distr = NULL, before) {
  if (is.null(distr)) {
    df_ma <-
      df_malaria_loreto %>%
      select(-soilm) %>%
      filter(province == prov) %>%
      group_by(province, dttm) %>%
      summarise(
        across(falciparum:vivax, ~ sum(., na.rm = TRUE)),
        across(aet:tmin, ~ mean(., na.rm = TRUE)),
        .groups = "drop"
      ) %>%
      arrange(dttm) %>%
      mutate(
        across(
          falciparum:tmin,
          ~ slide_index_dbl(., dttm, ~ mean(., na.rm = TRUE),
            .before = months(before), .complete = TRUE
          )
        )
      ) # %>% mutate(across(falciparum:tmin, ~ rescale(., to = c(0, 1))))
  } else {
    df_ma <-
      df_malaria_loreto %>%
      select(-soilm) %>%
      filter(province == prov, district == distr) %>%
      arrange(dttm) %>%
      mutate(
        across(
          falciparum:tmin,
          ~ slide_index_dbl(., dttm, ~ mean(., na.rm = TRUE),
            .before = months(before), .complete = TRUE
          )
        )
      ) # %>% mutate(across(falciparum:tmin, ~ rescale(., to = c(0, 1))))
  }

  df_ma_long <-
    df_ma %>%
    pivot_longer(
      cols = falciparum:tmin, names_to = "variable", values_to = "value"
    ) %>%
    mutate(
      variable = factor(
        variable,
        levels = c("falciparum", "vivax", "aet", "prcp", "q", "tmax", "tmin"),
        labels = c(
          "P. Falciparum", "P. Vivax", "Actual Evapotranspiration",
          "Precipitation", "Runoff", "Maximum Temperature",
          "Minimum Temperature"
        )
      )
    )

  plt_ma <- 
    df_ma_long %>%
    ggplot() +
    facet_wrap(vars(variable), ncol = 2, scales = "free_y") +
    geom_line(aes(dttm, value, color = variable)) +
    geom_rect(
      df_pamafro_period,
      mapping = aes(xmin = start, xmax = end),
      ymin = -Inf,
      ymax = Inf,
      alpha = 0.2
    ) +
    geom_vline(
      xintercept = c(df_pamafro_period$start, df_pamafro_period$end),
      linetype = "dashed",
      alpha = 0.6
    ) +
    scale_x_datetime(date_labels = "%Y", date_breaks = "2 year") +
    scale_color_npg() +
    scale_fill_npg() +
    theme(legend.position = "none")
    
  if (is.null(distr)) {
    plt <- plt_ma + labs(y = NULL, x = NULL, title = prov)
  } else {
    plt <- plt_ma + labs(y = NULL, x = NULL, title = paste0(prov, ", ", distr))
  }
  
  plt
}
```

```{r warning=FALSE}
plot_vars_ma(prov = "MAYNAS", before = 5)
```

```{r warning=FALSE}
provinces <- unique(df_malaria_loreto$province)
pdf("figs/provinces/provinces.pdf")
for (i in seq_along(provinces)) {
  print(plot_vars_ma(prov = provinces[i], before = 5))
}
dev.off()
```

```{r}
plot_vars_prov <- function(prov, before) {
  df_long <- 
    df_malaria_loreto %>% 
    select(-soilm) %>% 
    pivot_longer(aet:tmin, names_to = "variable", values_to = "value") %>% 
    mutate(
      variable = factor(
        variable,
        levels = c("aet", "prcp", "q", "tmax", "tmin"),
        labels = c(
          "Actual Evapotranspiration","Precipitation", "Runoff", 
          "Maximum Temperature", "Minimum Temperature"
        )
      )
    )
  
  df_ma <-
    df_malaria_loreto %>%
    select(-soilm) %>%
    filter(province == prov) %>%
    group_by(province, dttm) %>%
    summarise(across(aet:tmin, ~ mean(., na.rm = TRUE)), .groups = "drop") %>%
    arrange(dttm) %>%
    mutate(
      across(
        aet:tmin,
        ~ slide_index_dbl(., dttm, ~ mean(., na.rm = TRUE),
          .before = months(before), .complete = TRUE
        )
      )
    )

  df_ma_long <-
    df_ma %>%
    pivot_longer(aet:tmin, names_to = "variable", values_to = "value") %>%
    mutate(
      variable = factor(
        variable,
        levels = c("aet", "prcp", "q", "tmax", "tmin"),
        labels = c(
          "Actual Evapotranspiration","Precipitation", "Runoff", 
          "Maximum Temperature", "Minimum Temperature"
        )
      )
    )

  plt_ma <- 
    df_ma_long %>%
    ggplot() +
    facet_wrap(vars(variable), ncol = 2, scales = "free_y") +
    # geom_point(
    #   aes(dttm, value), df_long, color = "grey", size = 0.1, alpha = 0.5
    # ) +
    geom_line(
      aes(dttm, value, group = district), df_long, color = "grey", size = 0.1, alpha = 0.5
    ) +
    geom_line(aes(dttm, value, color = variable), size = 1) +
    geom_rect(
      aes(xmin = start, xmax = end),
      df_pamafro_period,
      ymin = -Inf,
      ymax = Inf,
      alpha = 0.2
    ) +
    geom_vline(
      xintercept = c(df_pamafro_period$start, df_pamafro_period$end),
      linetype = "dashed",
      alpha = 0.6
    ) +
    scale_x_datetime(date_labels = "%Y", date_breaks = "2 year") +
    scale_color_npg() +
    scale_fill_npg() +
    theme(legend.position = "none") + 
    labs(y = NULL, x = NULL, title = str_to_title(prov))
    
  plt_ma
}
```

```{r}
plot_vars_prov(prov = "MAYNAS", before = 5)
```

```{r}
for (i in seq_along(provinces)) {
  prov <- provinces[i] %>% str_replace_all(" ", "-") %>% tolower()
  png(
    paste0("figs/provinces/", prov, ".png"), width = 22, height = 18, units = "cm", 
    pointsize = 12, res = 300
  )
  print(plot_vars_prov(prov = provinces[i], before = 5))
  dev.off()
}
```


```{r warning=FALSE}
for (i in seq_along(provinces)) {
  districts <- 
    df_malaria_loreto %>% 
    filter(province == provinces[i]) %>% 
    select(district) %>%  
    unique() %>% 
    pull()
  prov <- provinces[i] %>% str_replace_all(" ", "-") %>% tolower()
  pdf(paste0("figs/districts/", prov, ".pdf"))
  for (j in seq_along(districts)) {
    print(plot_vars_ma(prov = provinces[i], distr = districts[j], before = 11))
  }
  dev.off()
}
```


```{r}
df_pamafro <- 
  df_malaria_loreto %>% 
  mutate(
    period = 
      ifelse(
        dttm < df_pamafro_period$start, "before", 
        ifelse(dttm > df_pamafro_period$end, "after", "during")
      )
  ) %>% 
  mutate(
    period = factor(period, levels = c("before", "during", "after"))
  ) %>% 
  mutate(
    after = ifelse(period == "after", "yes", "no")
  )
```

```{r}
df_pamafro %>% 
  select(-c(aet, soilm, q)) %>%
  # arrange(dttm) %>%
  # mutate(
  #   across(
  #     falciparum:tmin,
  #     ~ slide_index_dbl(
  #       ., dttm, ~ mean(., na.rm = TRUE), .before = months(5),
  #       .complete = TRUE
  #     )
  #   )
  # ) %>%
  group_by(province, district, after) %>% 
  summarise(
    across(prcp:tmin, ~ mean(., na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  group_by(province, after) %>% 
  summarise(
    across(prcp:tmin, ~ max(., na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  filter(province == "DATEM DEL MARAÑON")
```

```{r}
df_malaria_loreto %>% 
  group_by(province, district, year) %>% 
  summarise(
    vivax      = mean(vivax),
    falciparum = mean(falciparum),
    pop        = mean(pop2015),
    aet        = mean(aet),
    prcp       = mean(prcp),
    q          = mean(q),
    tmax       = mean(tmax),
    tmin       = mean(tmin),
    .groups = "drop"
  ) %>% 
  group_by(province, district) %>% 
  summarise(
    air_vivax      = 10000 * (mean(vivax) / mean(pop)),
    air_falciparum = 10000 * (mean(falciparum) / mean(pop)),
    aet            = mean(aet),
    prcp           = mean(prcp),
    q              = mean(q),
    tmax           = mean(tmax),
    tmin           = mean(tmin),
    pop = mean(pop),
    .groups = "drop"
  ) %>% 
  arrange(desc(air_vivax))
```


Year-over-year percentage change of the number of malaria cases reported (x1000) in Loreto from 2000 to 2017 by specie by specie.

```{r message=FALSE, warning=FALSE}
# dataset %>% 
#   dplyr::group_by(year) %>% 
#   dplyr::summarise(
#     falciparum = sum(falciparum),
#     vivax = sum(vivax)
#   ) %>% 
#   dplyr::arrange(year) %>% 
#   dplyr::mutate(
#     yoy_falciparum = (falciparum - lag(falciparum)) / lag(falciparum),
#     yoy_vivax      = (vivax - lag(vivax)) / lag(vivax)
#   ) %>% 
#   tidyr::pivot_longer(
#     cols     = c(yoy_falciparum, yoy_vivax),
#     names_to = "specie"
#   ) %>% 
#   dplyr::mutate(
#     specie = factor(
#       x      = specie, 
#       levels = c("yoy_falciparum", "yoy_vivax"), 
#       labels = c("P. falciparum", "P. vivax")
#     )
#   ) %>% 
#   ggplot(aes(x = year, y = value, fill = specie)) +
#   geom_col(position = "dodge") +
#   scale_color_manual(values = c("#00AFBB", "#E7B800")) +
#   geom_rect(
#     data    = tibble(start = 2005, end = 2010),
#     mapping = aes(xmin = start, xmax = end), 
#     ymin    = 0, 
#     ymax    = 100,
#     alpha   = 0.2
#   ) +
#   geom_vline(
#     data = tibble(dates = c(2005, 2010)),
#     aes(xintercept = dates),
#     linetype = "dashed",
#     colour   = "red",
#     alpha    = 0.8
#   ) +
#   geom_hline(
#     mapping  = aes(yintercept = 0),
#     linetype = "dashed",
#     alpha    = 0.6
#   ) +
#   scale_y_continuous(labels = scales::percent) +
#   labs(y = "Total number of malaria cases reported (x1000)", x = NULL) +
#   theme(legend.position = "top", legend.title = element_blank()) +
#   facet_wrap(~ specie)
```

```{r message=FALSE}
df_annual_summary %>% 
  dplyr::group_by(province) %>% 
  dplyr::summarise(
    mean_air_vivax      = round(10000 * mean(air_vivax), 2),
    sd_air_vivax        = round(sd(10000 * air_vivax), 2),
    mean_air_falciparum = round(10000 * mean(air_falciparum), 2),
    sd_air_falciparum   = round(sd(10000 * air_falciparum), 2),
    mean_aet            = round(mean(aet), 2),
    sd_aet              = round(sd(aet), 2),
    mean_prcp           = round(mean(prcp), 2),
    sd_prcp             = round(sd(prcp), 2),
    mean_q              = round(mean(q), 2),
    sd_q                = round(sd(q), 2),
    mean_tmax           = round(mean(tmax), 2),
    sd_tmax             = round(sd(tmax), 2),
    mean_tmin           = round(mean(tmin), 2),
    sd_tmin             = round(sd(tmin), 2)
  ) -> df_annual_summary_province 
```

```{r}
table_01_sketch = htmltools::withTags(
  table(
    class = "display",
    thead(
      tr(
        th(rowspan = 2, "Province"),
        th(colspan = 2, "AIR (P. vivax)"),
        th(colspan = 2, "IR (P. falciparum)"),
        th(colspan = 2, "Actual Evapotranspiration"),
        th(colspan = 2, "Precipitation"),
        th(colspan = 2, "Runoff"),
        th(colspan = 2, "Max Temperature"),
        th(colspan = 2, "Min Temperature")
      ),
      tr(
        lapply(rep(c("Mean", "SD"), 7), th)
      )
    )
  )
)
```

## Table 1

```{r}
df_annual_summary_province %>% 
  tibble::column_to_rownames(var = "province") %>% 
  DT::datatable(
    container  = table_01_sketch,
    extensions = 'Buttons',
    options    = list(
      dom        = 'Bt', 
      scrollX    = TRUE,
      columnDefs = list(list(className = 'dt-center', targets = 1:6)),
      buttons    = c("copy", "csv")
    )
  )
```

```{r message=FALSE}
df_annual_summary_v2 %>% 
  dplyr::group_by(province) %>%
  dplyr::summarise(
    mean_air_vivax      = round(10000 * mean(air_vivax), 2),
    sd_air_vivax        = round(sd(10000 * air_vivax), 2),
    mean_air_falciparum = round(10000 * mean(air_falciparum), 2),
    sd_air_falciparum   = round(sd(10000 * air_falciparum), 2),
    mean_aet            = round(mean(aet), 2),
    sd_aet              = round(sd(aet), 2),
    mean_prcp           = round(mean(prcp), 2),
    sd_prcp             = round(sd(prcp), 2),
    mean_q              = round(mean(q), 2),
    sd_q                = round(sd(q), 2),
    mean_tmax           = round(mean(tmax), 2),
    sd_tmax             = round(sd(tmax), 2),
    mean_tmin           = round(mean(tmin), 2),
    sd_tmin             = round(sd(tmin), 2)
  ) -> df_annual_summary_province_v2 
```

```{r}
table_01_sketch = htmltools::withTags(
  table(
    class = "display",
    thead(
      tr(
        th(rowspan = 2, "Province"),
        th(colspan = 2, "AIR (P. vivax)"),
        th(colspan = 2, "IR (P. falciparum)"),
        th(colspan = 2, "Actual Evapotranspiration"),
        th(colspan = 2, "Precipitation"),
        th(colspan = 2, "Runoff"),
        th(colspan = 2, "Max Temperature"),
        th(colspan = 2, "Min Temperature")
      ),
      tr(
        lapply(rep(c("Mean", "SD"), 7), th)
      )
    )
  )
)
```

## Table 1

```{r}
df_annual_summary_province_v2 %>% 
  tibble::column_to_rownames(var = "province") %>% 
  DT::datatable(
    container  = table_01_sketch,
    extensions = 'Buttons',
    options    = list(
      dom        = 'Bt', 
      scrollX    = TRUE,
      columnDefs = list(list(className = 'dt-center', targets = 1:6)),
      buttons    = c("copy", "csv")
    )
  )
```

# Correlation between climate variables

```{r eval=FALSE, include=FALSE}
# dataset %>% 
#   dplyr::group_by(year) %>% 
#   dplyr::summarise(
#     aet = mean(aet_std),
#     prcp = mean(prcp_std),
#     q = mean(q_std),
#     soilm = mean(soilm_std),
#     tmax = mean(tmax_std),
#     tmin = mean(tmin_std)
#   ) %>% 
#   dplyr::arrange(year) %>% 
#   tidyr::pivot_longer(
#     cols     = c(aet, prcp, q, soilm, tmax, tmin),
#     names_to = "variable"
#   ) %>% 
#   dplyr::mutate(
#     variable = factor(
#       x      = variable, 
#       levels = c("aet", "prcp", "q", "soilm", "tmax", "tmin"), 
#       labels = c(
#         "Actual Evapotranspiration", 
#         "Precipitation",
#         "Runoff",
#         "Soil Moisture",
#         "Max Temperature",
#         "Min Temperature"
#       )
#     )
#   ) -> variables_year_long
```

```{r eval=FALSE, include=FALSE}
# province_variables_year_plot <- 
#   ggplot(data = variables_year_long) +
#   geom_line(
#     mapping = aes(
#       x     = year, 
#       y     = value, 
#       color = variable, 
#       #fill  = variable
#     ),
#     size     = 1,
#     position = position_dodge(0)
#   ) +
#   geom_rect(
#     data    = tibble(start = 2005, end = 2010),
#     mapping = aes(xmin = start, xmax = end), 
#     ymin    = -100, 
#     ymax    = 100,
#     alpha   = 0.2
#   ) +
#   geom_vline(
#     xintercept = c(2005, 2010),
#     linetype = "dashed",
#     alpha = 0.8
#   ) +
#   geom_hline(
#     yintercept = 0,
#     linetype = "dashed",
#     alpha = 0.5
#   ) +
#   scale_y_continuous(limits = c(-1, 1)) +
#   scale_x_continuous(n.breaks = 16) +
#   labs(y = NULL, x = NULL) +
#   theme(legend.position = "top", legend.title = element_blank())
# 
# show(province_variables_year_plot)
```

```{r message=FALSE}
df_malaria_loreto %>%
  dplyr::filter(vivax > 0, falciparum > 0, district == "ALTO NANAY") %>% 
  dplyr::mutate(log_vivax = log(vivax), log_falciparum = log(falciparum)) %>% 
  dplyr::select(vivax, falciparum, aet, prcp, q, tmax, tmin) %>% 
  GGally::ggpairs() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r warning=FALSE}
rmc_aet_prcp <- rmcorr::rmcorr("district", "aet", "prcp", df_malaria_loreto)
print(rmc_aet_prcp$r)
```

```{r}
df_malaria_loreto %>% 
  ggplot(aes(x = aet, y = prcp, group = district, color = district)) +
  geom_point(aes(colour = district)) +
  geom_line(aes(y = rmc_aet_prcp$model$fitted.values), linetype = 1) +
  theme(legend.position = "none")
```

```{r warning=FALSE}
rmc_q_prcp <- rmcorr::rmcorr("district", "q", "prcp", df_malaria_loreto)
print(rmc_q_prcp$CI)
```

```{r}
df_malaria_loreto %>% 
  ggplot(aes(x = q, y = prcp, group = district, color = district)) +
  geom_point(aes(colour = district)) +
  geom_line(aes(y = rmc_q_prcp$model$fitted.values), linetype = 1) +
  theme(legend.position = "none")
```

```{r warning=FALSE}
covariates <- c("aet", "prcp", "q", "tmax", "tmin")

cor_matrix <- matrix(0, nrow = length(covariates), ncol = length(covariates))

for(i in 1:length(covariates)) {
  for(j in 1:length(covariates)) {
    cor_matrix[i, j] <- rmcorr::rmcorr(
      "district", 
      covariates[i], 
      covariates[j], 
      df_malaria_loreto
    )$r
  }
}
```

```{r}
rownames(cor_matrix) <- covariates
colnames(cor_matrix) <- covariates
cor_matrix
```

```{r warning=FALSE}
cor_lb_matrix <- matrix(0, nrow = length(covariates), ncol = length(covariates))

for(i in 1:length(covariates)) {
  for(j in 1:length(covariates)) {
    cor_lb_matrix[i, j] <- rmcorr::rmcorr(
      "district", 
      covariates[i], 
      covariates[j], 
      df_malaria_loreto
    )$CI[1]
  }
}
```

```{r}
rownames(cor_lb_matrix) <- covariates
colnames(cor_lb_matrix) <- covariates
cor_lb_matrix
```

```{r warning=FALSE}
cor_ub_matrix <- matrix(0, nrow = length(covariates), ncol = length(covariates))

for(i in 1:length(covariates)) {
  for(j in 1:length(covariates)) {
    cor_ub_matrix[i, j] <- rmcorr::rmcorr(
      "district", 
      covariates[i], 
      covariates[j], 
      df_malaria_loreto
    )$CI[2]
  }
}
```

```{r}
rownames(cor_ub_matrix) <- covariates
colnames(cor_ub_matrix) <- covariates
cor_ub_matrix
```
