---
title: "Exploratory data analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    fig_width: 12
    fig_height: 10
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---

```{r message=FALSE, warning=FALSE}
library(readr, warn.conflicts = FALSE)
library(skimr)

library(tibble)
library(lubridate, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)

library(ggplot2, warn.conflicts = FALSE)
ggplot2::theme_set(theme_bw())
ggplot2::theme_update(panel.grid = element_blank())

library(geofacet)
library(ggsci)

library(qwraps2)
options(qwraps2_markup = "markdown")

library(DT)
library(htmltools)
library(stringr)

```

# Data loading 

Create csv. file path. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
processed_data_path <- file.path("../data/processed/") 
file_name           <- "pro_malaria-monthly-cases-district-loreto"
file                <- paste(file_name, ".csv", sep = "")
file_path           <- file.path(processed_data_path, file) 
```

Define data types for columns.

```{r message=FALSE, warning=FALSE}
# Column names
col_names <- c(
  "district",
  "year",
  "month",
  "falciparum",
  "vivax",
  "aet",
  "prcp",
  "q",
  "soilm",
  "tmax",
  "tmin",
  "pop2015",
  "province",
  "region",
  "dttm",
  "time"
)
```

```{r}
# Column types
col_types <- 
  readr::cols(
    district      = col_character(),
    year          = col_integer(),
    month         = col_integer(),
    falciparum    = col_integer(),
    vivax         = col_integer(),
    aet           = col_double(),
    prcp          = col_double(),
    q             = col_double(),
    soilm         = col_double(),
    tmax          = col_double(),
    tmin          = col_double(),
    pop2015       = col_integer(),
    province      = col_character(),
    region        = col_character(),
    dttm          = col_datetime(),
    time          = col_integer()
  )
```

Reading csv. file.

```{r message=FALSE, warning=FALSE}
df_malaria_loreto <- 
  readr::read_csv(
    file      = file_path,
    col_names = col_names,
    col_types = col_types,
    skip      = 1,
    locale    = locale(encoding = "UTF-8")
  )
```

## Data validation 

TODO: Unique values evaluation.

Check data structure.

```{r message=FALSE, warning=FALSE}
str(df_malaria_loreto)
```

Summarize

```{r summary}
skimr::skim(df_malaria_loreto)
```

Check data structure

```{r data-structure}
str(df_malaria_loreto)
```

Create a data frame with the start and end dates of PAMAFRO intervention.

```{r message=FALSE, warning=FALSE}
df_pamafro_period <- tibble::tibble(
  start = lubridate::make_datetime(year = 2005, month = 10, day = 1L), 
  end   = lubridate::make_datetime(year = 2010, month = 09, day = 1L)
)
```

# Exploratory data analysis

TODO: Summarize by districts. `soilm` have the same value for all months in some districts. Do not include this variable in the analysis.

```{r message=FALSE}
df_malaria_loreto %>% 
  dplyr::group_by(province, year, month) %>% 
  dplyr::summarise(
    vivax      = sum(vivax),
    falciparum = sum(falciparum),
    pop        = sum(pop2015),
    aet        = mean(aet),
    prcp       = mean(prcp),
    q          = mean(q),
    tmax       = mean(tmax),
    tmin       = mean(tmin)
  ) %>% 
  dplyr::mutate(
    ir_vivax      = vivax / pop,
    ir_falciparum = falciparum / pop
  ) %>% 
  dplyr::group_by(province, year) %>% 
  dplyr::summarise(
    air_vivax      = mean(ir_vivax),
    air_falciparum = mean(ir_falciparum),
    aet            = mean(aet),
    prcp           = mean(prcp),
    q              = mean(q),
    tmax           = mean(tmax),
    tmin           = mean(tmin)
  ) -> df_annual_summary
```

```{r}
df_annual_summary %>% 
  tidyr::pivot_longer(
    cols      = c(ir_falciparum, ir_vivax),
    names_to  = "specie",
    values_to = "ir"
  ) %>% 
  dplyr::mutate(
    specie = factor(
      x      = specie, 
      levels = c("ir_falciparum", "ir_vivax"), 
      labels = c("P. falciparum", "P. vivax")
    )
  ) -> df_annual_summary_long
```

```{r}
# geofacet::grid_design()
```

```{r}
loreto_grid <- data.frame(
  row  = c(2, 3, 3, 3, 3, 4, 4, 5) - 1,
  col  = c(3, 2, 4, 1, 3, 2, 3, 2),
  code = c("MA", "LO", "PU", "DM", "RC", "AA", "RE", "UC"),
  name = c(
    "MAYNAS", 
    "LORETO", 
    "PUTUMAYO", 
    "DATEM DEL MARAÑON", 
    "MARISCAL RAMON CASTILLA", 
    "ALTO AMAZONAS", 
    "REQUENA", 
    "UCAYALI"
  ),
  stringsAsFactors = FALSE
)
```

## Figure 1

```{r message=FALSE, fig.height=8, fig.width=12}
plot_air_province <- 
  ggplot(data = df_annual_summary_long) +
  geofacet::facet_geo(~ province, grid = loreto_grid, label = "name") +
  geom_area(
    mapping  = aes(x = year, y = ir, color = specie, fill = specie),
    size     = 1,
    alpha    = 0.3,
    position = position_dodge(0)
  ) +
  geom_rect(
    data = tibble::tibble(start = 2005, end = 2010),
    mapping = aes(xmin = start, xmax = end), 
    ymin    = -100, 
    ymax    = 100,
    alpha   = 0.2
  ) +
  geom_vline(xintercept = c(2005, 2010), linetype = "dashed", alpha = 0.6) +
  scale_y_continuous(limits = c(0, 15)) +
  scale_x_continuous(n.breaks = 5) +
  ggsci::scale_color_npg() +
  ggsci::scale_fill_npg() +
  labs(y = "Annual incidence rate", x = NULL) +
  theme(legend.position = "top", legend.title = element_blank())

show(plot_air_province)
```

Year-over-year percentage change of the number of malaria cases reported (x1000) in Loreto from 2000 to 2017 by specie by specie.

```{r message=FALSE, warning=FALSE}
# dataset %>% 
#   dplyr::group_by(year) %>% 
#   dplyr::summarise(
#     falciparum = sum(falciparum),
#     vivax = sum(vivax)
#   ) %>% 
#   dplyr::arrange(year) %>% 
#   dplyr::mutate(
#     yoy_falciparum = (falciparum - lag(falciparum)) / lag(falciparum),
#     yoy_vivax      = (vivax - lag(vivax)) / lag(vivax)
#   ) %>% 
#   tidyr::pivot_longer(
#     cols     = c(yoy_falciparum, yoy_vivax),
#     names_to = "specie"
#   ) %>% 
#   dplyr::mutate(
#     specie = factor(
#       x      = specie, 
#       levels = c("yoy_falciparum", "yoy_vivax"), 
#       labels = c("P. falciparum", "P. vivax")
#     )
#   ) %>% 
#   ggplot(aes(x = year, y = value, fill = specie)) +
#   geom_col(position = "dodge") +
#   scale_color_manual(values = c("#00AFBB", "#E7B800")) +
#   geom_rect(
#     data    = tibble(start = 2005, end = 2010),
#     mapping = aes(xmin = start, xmax = end), 
#     ymin    = 0, 
#     ymax    = 100,
#     alpha   = 0.2
#   ) +
#   geom_vline(
#     data = tibble(dates = c(2005, 2010)),
#     aes(xintercept = dates),
#     linetype = "dashed",
#     colour   = "red",
#     alpha    = 0.8
#   ) +
#   geom_hline(
#     mapping  = aes(yintercept = 0),
#     linetype = "dashed",
#     alpha    = 0.6
#   ) +
#   scale_y_continuous(labels = scales::percent) +
#   labs(y = "Total number of malaria cases reported (x1000)", x = NULL) +
#   theme(legend.position = "top", legend.title = element_blank()) +
#   facet_wrap(~ specie)
```

```{r message=FALSE}
df_annual_summary %>% 
  dplyr::group_by(province) %>% 
  dplyr::summarise(
    mean_air_vivax      = round(10000 * mean(air_vivax), 2),
    sd_air_vivax        = round(sd(10000 * air_vivax), 2),
    mean_air_falciparum = round(10000 * mean(air_falciparum), 2),
    sd_air_falciparum   = round(sd(10000 * air_falciparum), 2),
    mean_aet            = round(mean(aet), 2),
    sd_aet              = round(sd(aet), 2),
    mean_prcp           = round(mean(prcp), 2),
    sd_prcp             = round(sd(prcp), 2),
    mean_q              = round(mean(q), 2),
    sd_q                = round(sd(q), 2),
    mean_tmax           = round(mean(tmax), 2),
    sd_tmax             = round(sd(tmax), 2),
    mean_tmin           = round(mean(tmin), 2),
    sd_tmin             = round(sd(tmin), 2)
  ) -> df_annual_summary_province 
```

```{r}
table_01_sketch = htmltools::withTags(
  table(
    class = "display",
    thead(
      tr(
        th(rowspan = 2, "Province"),
        th(colspan = 2, "AIR (P. vivax)"),
        th(colspan = 2, "IR (P. falciparum)"),
        th(colspan = 2, "Actual Evapotranspiration"),
        th(colspan = 2, "Precipitation"),
        th(colspan = 2, "Runoff"),
        th(colspan = 2, "Max Temperature"),
        th(colspan = 2, "Min Temperature")
      ),
      tr(
        lapply(rep(c("Mean", "SD"), 7), th)
      )
    )
  )
)
```

## Table 1

```{r}
df_annual_summary_province %>% 
  tibble::column_to_rownames(var = "province") %>% 
  DT::datatable(
    container  = table_01_sketch,
    extensions = 'Buttons',
    options    = list(
      dom        = 'Bt', 
      scrollX    = TRUE,
      columnDefs = list(list(className = 'dt-center', targets = 1:6)),
      buttons    = c("copy", "csv")
    )
  )
```

Climate variables over time

```{r}
# dataset %>% 
#   dplyr::group_by(year) %>% 
#   dplyr::summarise(
#     aet = mean(aet_std),
#     prcp = mean(prcp_std),
#     q = mean(q_std),
#     soilm = mean(soilm_std),
#     tmax = mean(tmax_std),
#     tmin = mean(tmin_std)
#   ) %>% 
#   dplyr::arrange(year) %>% 
#   tidyr::pivot_longer(
#     cols     = c(aet, prcp, q, soilm, tmax, tmin),
#     names_to = "variable"
#   ) %>% 
#   dplyr::mutate(
#     variable = factor(
#       x      = variable, 
#       levels = c("aet", "prcp", "q", "soilm", "tmax", "tmin"), 
#       labels = c(
#         "Actual Evapotranspiration", 
#         "Precipitation",
#         "Runoff",
#         "Soil Moisture",
#         "Max Temperature",
#         "Min Temperature"
#       )
#     )
#   ) -> variables_year_long
```

```{r}
# province_variables_year_plot <- 
#   ggplot(data = variables_year_long) +
#   geom_line(
#     mapping = aes(
#       x     = year, 
#       y     = value, 
#       color = variable, 
#       #fill  = variable
#     ),
#     size     = 1,
#     position = position_dodge(0)
#   ) +
#   geom_rect(
#     data    = tibble(start = 2005, end = 2010),
#     mapping = aes(xmin = start, xmax = end), 
#     ymin    = -100, 
#     ymax    = 100,
#     alpha   = 0.2
#   ) +
#   geom_vline(
#     xintercept = c(2005, 2010),
#     linetype = "dashed",
#     alpha = 0.8
#   ) +
#   geom_hline(
#     yintercept = 0,
#     linetype = "dashed",
#     alpha = 0.5
#   ) +
#   scale_y_continuous(limits = c(-1, 1)) +
#   scale_x_continuous(n.breaks = 16) +
#   labs(y = NULL, x = NULL) +
#   theme(legend.position = "top", legend.title = element_blank())
# 
# show(province_variables_year_plot)
```
