---
title: "Time-varying coefficients model using GAMs"
output: 
  html_document:
    toc: true
editor_options: 
  chunk_output_type: inline
---

We are going to use *generalized additive models* (GAMs) to estimate time-varying coefficients for climate variables in order to quantify their effects on malaria incidence by specie thought the time period of the study.

First, lets load some packages to handle the tasks we are going to do.

```{r message=FALSE, warning=FALSE}
# Utilities
library(tictoc)

# Data loading
library(readr)
library(skimr)

# Data manipulation
library(tibble)
library(lubridate)
library(dplyr)
library(splitTools)
library(scales)
library(DT)
library(htmltools)
library(stringr)

# Plots
library(ggplot2)
ggplot2::theme_set(theme_bw())
ggplot2::theme_update(panel.grid = element_blank())
library(ggsci)

# GAM
library(mgcv)
library(gratia)
```

# Data wrangling

## Loading

The cleaned up data can be found on the `~/data/processed/` directory on this repository. Define the file path to read the data. 

```{r}
processed_data_path <- file.path("../data/processed/") 
file_name           <- "pro_malaria-monthly-cases-district-loreto"
file                <- paste(file_name, ".csv", sep = "")
file_path           <- file.path(processed_data_path, file) 
```

Create a variable containing the column names for the data.

```{r message=FALSE, warning=FALSE}
col_names <- c(
  "district",
  "year",
  "month",
  "falciparum",
  "vivax",
  "aet",
  "prcp",
  "q",
  "soilm",
  "tmax",
  "tmin",
  "pop2015",
  "province",
  "region",
  "dttm",
  "time"
)
```

Define column types to optimize the parsing. We exclude `soilm` and `region` for the analysis.

```{r}
# Column types
col_types <- 
  readr::cols_only(
    district      = col_character(),
    year          = col_integer(),
    month         = col_integer(),
    falciparum    = col_integer(),
    vivax         = col_integer(),
    aet           = col_double(),
    prcp          = col_double(),
    q             = col_double(),
    tmax          = col_double(),
    tmin          = col_double(),
    pop2015       = col_integer(),
    province      = col_character(),
    dttm          = col_datetime(),
    time          = col_integer()
  )
```

Reading csv. file into a data frame.

```{r message=FALSE, warning=FALSE}
df_malaria_loreto <- 
  readr::read_csv(
    file      = file_path,
    col_names = col_names,
    col_types = col_types,
    skip      = 1,
    locale    = readr::locale(encoding = "UTF-8")
  )
```

## Parsing

We are going to parse `district`, `province`, `year` and `month` as factors.

```{r}
to_factor <- c("year", "month", "district", "province")

for (col in to_factor) {
  df_malaria_loreto[[col]] <- factor(df_malaria_loreto[[col]])
}
```

## Validation 

Check data structure.

```{r message=FALSE, warning=FALSE}
str(df_malaria_loreto)
```

To display a summary of the data, we first define a variable containing a table with summary statistics for each data type in the data frame using the `skim` function in the `skimr` package.

```{r}
skim_malaria <- skimr::skim(df_malaria_loreto)
```


Then, we display the summary tables for each data type using the `yank` function of the same package:

- Factors:

```{r}
skimr::yank(skim_malaria, "factor")
```

- Numeric:

```{r}
skimr::yank(skim_malaria, "numeric")
```

- Datetime:

```{r}
skimr::yank(skim_malaria, "POSIXct")
```

## Feature engineering

Create a copy of the data frame where we are going to transform and create some new columns.

```{r}
dat <- tibble::tibble(df_malaria_loreto)
```

### Scaling

First, we are going to scale all climate variables from 0 to 1. The aim of this step is to avoid numerical problems in the GAM fitting procedures, and also to have a common scale on the coefficients.

```{r}
to_scale <- c("aet", "prcp", "q", "soilm", "tmax", "tmin")

for (col in to_scale) {
  dat[[col]] <- scales::rescale(df_malaria_loreto[[col]], to = c(0, 1))
}
```

### Lagged variables

Lagged variables are going to be used to build *distributed lag models* (DLM). In order to properly create lagged variables for this data set, first we are going to group the data by districts. Next, we arrange the data in each group by date. Finally, la `lag` function of the package `dplyr` is applied to create lagged variables by 1, 3, 6, and 12 months.

```{r}
dat <- dat %>% 
  dplyr::group_by(district) %>%
  dplyr::arrange(dttm) %>% 
  dplyr::mutate(
    aet_lag1    = dplyr::lag(aet,  n = 1L),
    aet_lag3    = dplyr::lag(aet,  n = 3L),
    aet_lag6    = dplyr::lag(aet,  n = 6L),
    aet_lag12   = dplyr::lag(aet,  n = 12L),
    prcp_lag1   = dplyr::lag(prcp, n = 1L),
    prcp_lag3   = dplyr::lag(prcp, n = 3L),
    prcp_lag6   = dplyr::lag(prcp, n = 6L),
    prcp_lag12  = dplyr::lag(prcp, n = 12L),
    q_lag1      = dplyr::lag(q,    n = 1L),
    q_lag3      = dplyr::lag(q,    n = 3L),
    q_lag6      = dplyr::lag(q,    n = 6L),
    q_lag12     = dplyr::lag(q,    n = 12L),
    soilm_lag1  = dplyr::lag(q,    n = 1L),
    soilm_lag3  = dplyr::lag(q,    n = 3L),
    soilm_lag6  = dplyr::lag(q,    n = 6L),
    soilm_lag12 = dplyr::lag(q,    n = 12L),
    tmax_lag1   = dplyr::lag(tmax, n = 1L),
    tmax_lag3   = dplyr::lag(tmax, n = 3L),
    tmax_lag6   = dplyr::lag(tmax, n = 6L),
    tmax_lag12  = dplyr::lag(tmax, n = 12L),
    tmin_lag1   = dplyr::lag(tmin, n = 1L),
    tmin_lag3   = dplyr::lag(tmin, n = 3L),
    tmin_lag6   = dplyr::lag(tmin, n = 6L),
    tmin_lag12  = dplyr::lag(tmin, n = 12L)
  )
```

Since `NA` values are going to be present for the first 12 months (because of the 12 month lagged variables), we have to do a filter.

```{r}
dat <- na.omit(dat)
```

Finally, we create a column containing the values of the districts' population on the natural logarithmic scale.

```{r}
dat$ln_pop2015 <- log(dat$pop2015)
```

Display the final data frame for checking.

```{r}
dat
```
# Model building

Mainly, we are going to use the *Akaike information criterion* (AIC) to compare the candidate models. The AIC measures how well a model fits the data given a certain number of parameters in the model. For further examination of the models, we are going to extract a random test set of the data. The aim is to see how well the model interpolates unseen observations during the time period of the study. We are going to use the out-of-bag deviance to measure this model feature.

```{r}
set.seed(20210223)

indices <- splitTools::partition(
  y    = dat$district, 
  type = "stratified", 
  p    = c(train = 0.88, test = 0.12)
)

train <- dat[indices$train, ]
test  <- dat[indices$test, ]
```

Define function that returns the out-of-sample deviance for some model.

```{r}
# TODO Cite authors
get_oos_dev <- function(model, prediction, y, weights = NULL) {
  
  if(is.null(weights)) weights <- rep(1, times= length(y))
  
  dev_residuals <- model$family$dev.resids(y, prediction, weights)
  deviance      <- sum(dev_residuals)
  deviance
}
```

## TVCM: Model 0

Predictors: `aet`, `prcp`, `q`, `tmax`, `tmin`

### P. vivax

```{r}
tictoc::tic("GAM model fitting")
vivax_tvcm_0 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,   bs = "tp", k = 50) +
    s(time, by = prcp,  bs = "tp", k = 30) +
    s(time, by = q,     bs = "tp", k = 30) +
    s(time, by = tmax,  bs = "tp", k = 100) +
    s(time, by = tmin,  bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_0$converged, "\n")

mgcv::summary.gam(vivax_tvcm_0)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_0, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_0, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_0)
```

```{r}
mgcv::concurvity(vivax_tvcm_0, full = FALSE)$estimate
```

### P. falciparum

```{r}
tictoc::tic("GAM model fitting")
falciparum_tvcm_0 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,   bs = "tp", k = 50) +
    s(time, by = prcp,  bs = "tp", k = 30) +
    s(time, by = q,     bs = "tp", k = 30) +
    s(time, by = tmax,  bs = "tp", k = 100) +
    s(time, by = tmin,  bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_0$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_0)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_0, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_0, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_0)
```

```{r}
mgcv::concurvity(falciparum_tvcm_0, full = FALSE)$estimate
```

## TVCM: Model 1

Predictors: `aet`, `q`, `tmax`, `tmin`

### P. vivax

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_1 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,   bs = "tp", k = 50) +
    s(time, by = q,     bs = "tp", k = 30) +
    s(time, by = tmax,  bs = "tp", k = 100) +
    s(time, by = tmin,  bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_1$converged, "\n")

mgcv::summary.gam(vivax_tvcm_1)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_1, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_1, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_1)
```

```{r}
mgcv::concurvity(vivax_tvcm_1, full = FALSE)$estimate
```

### P. falciparum

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_1 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,   bs = "tp", k = 50) +
    s(time, by = q,     bs = "tp", k = 30) +
    s(time, by = tmax,  bs = "tp", k = 100) +
    s(time, by = tmin,  bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_1$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_1)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_1, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_1, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_1)
```

```{r}
mgcv::concurvity(falciparum_tvcm_1, full = FALSE)$estimate
```

## TVCM: Model 2

Predictors: `aet`, `prcp`, `tmax`, `tmin`

### P. vivax

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_2 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,   bs = "tp", k = 50) +
    s(time, by = prcp,  bs = "tp", k = 30) +
    s(time, by = tmax,  bs = "tp", k = 100) +
    s(time, by = tmin,  bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_2$converged, "\n")

mgcv::summary.gam(vivax_tvcm_2)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_2, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_2, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_2)
```

```{r}
mgcv::concurvity(vivax_tvcm_2, full = FALSE)$estimate
```

### P. falciparum

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_2 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,   bs = "tp", k = 50) +
    s(time, by = prcp,  bs = "tp", k = 30) +
    s(time, by = tmax,  bs = "tp", k = 100) +
    s(time, by = tmin,  bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_2$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_2)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_2, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_2, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_2)
```

```{r}
mgcv::concurvity(falciparum_tvcm_2, full = FALSE)$estimate
```

## TVCM: Model 3

### P. vivax

Predictors: `aet`, `q`, `tmin`

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_3 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,  bs = "tp", k = 50) +
    s(time, by = q,    bs = "tp", k = 30) +
    s(time, by = tmin, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_3$converged, "\n")

mgcv::summary.gam(vivax_tvcm_3)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_3, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_3, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_3)
```

```{r}
mgcv::concurvity(vivax_tvcm_3, full = FALSE)$estimate
```

### P. falciparum

Predictors: `aet`, `q`, `tmin`

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_3 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,  bs = "tp", k = 50) +
    s(time, by = q,    bs = "tp", k = 30) +
    s(time, by = tmin, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_3$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_3)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_3, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_3, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_3)
```

```{r}
mgcv::concurvity(falciparum_tvcm_3, full = FALSE)$estimate
```

## TVCM: Model 4

Predictors: `aet`, `q`, `tmax`

### P. vivax

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_4 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,  bs = "tp", k = 50) +
    s(time, by = q,    bs = "tp", k = 30) +
    s(time, by = tmax, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_4$converged, "\n")

mgcv::summary.gam(vivax_tvcm_4)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_4, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_4, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_4)
```

```{r}
mgcv::concurvity(vivax_tvcm_4, full = FALSE)$estimate
```

### P. falciparum

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_4 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,  bs = "tp", k = 50) +
    s(time, by = q,    bs = "tp", k = 30) +
    s(time, by = tmax, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_4$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_4)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_4, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_4, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_4)
```

```{r}
mgcv::concurvity(falciparum_tvcm_4, full = FALSE)$estimate
```

## TVCM: Model 5

Predictors: `aet`, `prcp`, `tmin`

### P. vivax

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_5 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,  bs = "tp", k = 50) +
    s(time, by = prcp, bs = "tp", k = 30) +
    s(time, by = tmin, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_5$converged, "\n")

mgcv::summary.gam(vivax_tvcm_5)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_5, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_5, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_5)
```

```{r}
mgcv::concurvity(vivax_tvcm_5, full = FALSE)$estimate
```

### P. falciparum

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_5 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,  bs = "tp", k = 50) +
    s(time, by = prcp, bs = "tp", k = 30) +
    s(time, by = tmin, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_5$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_5)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_5, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_5, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_5)
```

```{r}
mgcv::concurvity(falciparum_tvcm_5, full = FALSE)$estimate
```

## TVCM: Model 6

Predictors: `aet`, `prcp`,`tmax`

### P. vivax

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_6 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,  bs = "tp", k = 50) +
    s(time, by = prcp, bs = "tp", k = 30) +
    s(time, by = tmax, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_6$converged, "\n")

mgcv::summary.gam(vivax_tvcm_6)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_6, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_6, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_6)
```

```{r}
mgcv::concurvity(vivax_tvcm_6, full = FALSE)$estimate
```

### P. falciparum

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_6 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = aet,  bs = "tp", k = 50) +
    s(time, by = prcp, bs = "tp", k = 30) +
    s(time, by = tmax, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_6$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_6)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_6, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_6, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_6)
```

```{r}
mgcv::concurvity(falciparum_tvcm_6, full = FALSE)$estimate
```

## TVCM: Model 7

Predictors: `q`, `tmin`

### P. vivax

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_7 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = q,    bs = "tp", k = 30) +
    s(time, by = tmin, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_7$converged, "\n")

mgcv::summary.gam(vivax_tvcm_7)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_7, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_7, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_7)
```

```{r}
mgcv::concurvity(vivax_tvcm_7, full = FALSE)$estimate
```

### P. falciparum

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_7 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = q,    bs = "tp", k = 30) +
    s(time, by = tmin, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_7$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_7)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_7, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_7, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_7)
```

```{r}
mgcv::concurvity(falciparum_tvcm_7, full = FALSE)$estimate
```

## TVCM: Model 8

Predictors: `q`, `tmax`

### P. vivax

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_8 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = q,    bs = "tp", k = 30) +
    s(time, by = tmax, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_8$converged, "\n")

mgcv::summary.gam(vivax_tvcm_8)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_8, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_8, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_8)
```

```{r}
mgcv::concurvity(vivax_tvcm_8, full = FALSE)$estimate
```

### P. falciparum

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_8 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = q,    bs = "tp", k = 30) +
    s(time, by = tmax, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_8$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_8)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_8, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_8, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_8)
```

```{r}
mgcv::concurvity(falciparum_tvcm_8, full = FALSE)$estimate
```

## TVCM: Model 9

Predictors: `prcp`, `tmin`

### P. vivax 

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_9 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = prcp, bs = "tp", k = 30) +
    s(time, by = tmin, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_9$converged, "\n")

mgcv::summary.gam(vivax_tvcm_9)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_9, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_9, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_9)
```

```{r}
mgcv::concurvity(vivax_tvcm_9, full = FALSE)$estimate
```

### P. falciparum 

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_9 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = prcp, bs = "tp", k = 30) +
    s(time, by = tmin, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_9$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_9)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_9, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_9, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_9)
```

```{r}
mgcv::concurvity(falciparum_tvcm_9, full = FALSE)$estimate
```

## TVCM: Model 10

Predictors: `prcp`, `tmax`

### P. vivax

```{r message=FALSE}
tictoc::tic("GAM model fitting")
vivax_tvcm_10 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = prcp, bs = "tp", k = 30) +
    s(time, by = tmax, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvcm_10$converged, "\n")

mgcv::summary.gam(vivax_tvcm_10)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(vivax_tvcm_10, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = vivax_tvcm_10, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(vivax_tvcm_10)
```

```{r}
mgcv::concurvity(vivax_tvcm_10, full = FALSE)$estimate
```

### P. falciparum

```{r message=FALSE}
tictoc::tic("GAM model fitting")
falciparum_tvcm_10 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) +
    s(district, bs = "re", k = 49) +
    s(time, by = prcp, bs = "tp", k = 30) +
    s(time, by = tmax, bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", falciparum_tvcm_10$converged, "\n")

mgcv::summary.gam(falciparum_tvcm_10)
```

```{r message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
mgcv::gam.check(falciparum_tvcm_10, pch = 19, cex = 0.3)
```

```{r message=FALSE, warning=FALSE}
mgcv::plot.gam(
  x          = falciparum_tvcm_10, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0, 
  pages      = 1
)
```

```{r}
mgcv::concurvity(falciparum_tvcm_10)
```

```{r}
mgcv::concurvity(falciparum_tvcm_10, full = FALSE)$estimate
```

### Model selection

#### P. vivax

```{r}
vivax_tvc_models <- list(
  vivax_tvcm_0,
  vivax_tvcm_1,
  vivax_tvcm_2,
  vivax_tvcm_3,
  vivax_tvcm_4,
  vivax_tvcm_5,
  vivax_tvcm_6,
  vivax_tvcm_7,
  vivax_tvcm_8,
  vivax_tvcm_9,
  vivax_tvcm_10
)

n_vivax_tvc_models <- length(vivax_tvc_models)

aic_tvcm_vivax <- rep(NA, n_vivax_tvc_models)
oos_dev_tvcm_vivax <- rep(NA, n_vivax_tvc_models)
preds_tvcm_vivax <- vector(mode = "list", length = n_vivax_tvc_models)
names_tvcm_vivax <- rep(NA, n_vivax_tvc_models)

for(i in 1:n_vivax_tvc_models) {
  aic_tvcm_vivax[i] <- vivax_tvc_models[[i]]$aic
  preds_tvcm_vivax[[i]] <- mgcv::predict.bam(
    vivax_tvc_models[[i]],
    test,
    type = "response"
  )
  oos_dev_tvcm_vivax[i] <- get_oos_dev(
    vivax_tvc_models[[i]],
    preds_tvcm_vivax[[i]],
    test$vivax
  )
  
  names_tvcm_vivax[i] <- paste("Model", (i-1))
}
```

```{r}
eval_vivax_tvcm <- tibble::tibble(
  model   = names_tvcm_vivax,
  aic     = aic_tvcm_vivax,
  oos_dev = oos_dev_tvcm_vivax
)
```

```{r}
eval_vivax_tvcm %>% dplyr::arrange(aic)
```
```{r}
eval_vivax_tvcm %>% dplyr::arrange(oos_dev)
```
#### P. falciparum

```{r}
falciparum_tvc_models <- list(
  falciparum_tvcm_0,
  falciparum_tvcm_1,
  falciparum_tvcm_2,
  falciparum_tvcm_3,
  falciparum_tvcm_4,
  falciparum_tvcm_5,
  falciparum_tvcm_6,
  falciparum_tvcm_7,
  falciparum_tvcm_8,
  falciparum_tvcm_9,
  falciparum_tvcm_10
)

n_falciparum_tvc_models <- length(falciparum_tvc_models)

aic_tvcm_falciparum <- rep(NA, n_falciparum_tvc_models)
oos_dev_tvcm_falciparum <- rep(NA, n_falciparum_tvc_models)
preds_tvcm_falciparum <- vector(mode = "list", length = n_falciparum_tvc_models)
names_tvcm_falciparum <- rep(NA, n_falciparum_tvc_models)

for(i in 1:n_falciparum_tvc_models) {
  aic_tvcm_falciparum[i] <- falciparum_tvc_models[[i]]$aic
  preds_tvcm_falciparum[[i]] <- mgcv::predict.bam(
    falciparum_tvc_models[[i]],
    test,
    type = "response"
  )
  oos_dev_tvcm_falciparum[i] <- get_oos_dev(
    falciparum_tvc_models[[i]],
    preds_tvcm_falciparum[[i]],
    test$falciparum
  )
  
  names_tvcm_falciparum[i] <- paste("Model", (i-1))
}
```

```{r}
eval_falciparum_tvcm <- tibble::tibble(
  model   = names_tvcm_falciparum,
  aic     = aic_tvcm_falciparum,
  oos_dev = oos_dev_tvcm_falciparum
)
```

```{r}
eval_falciparum_tvcm %>% dplyr::arrange(aic)
```
```{r}
eval_falciparum_tvcm %>% dplyr::arrange(oos_dev)
```

## Figure 2

```{r message=FALSE, warning=FALSE}
df_pamafro_period <- tibble::tibble(
  start = lubridate::make_datetime(year = 2005, month = 10, day = 1L), 
  end   = lubridate::make_datetime(year = 2010, month = 09, day = 1L)
)
```

```{r include=FALSE}
plot_vivax_tvcm_5 <- mgcv::plot.gam(
  x          = vivax_tvcm_5,
  n          = 500,
  seWithMean = TRUE, 
  scale      = 0
)

plot_falciparum_tvcm_5 <- mgcv::plot.gam(
  x          = falciparum_tvcm_5,
  n          = 500,
  seWithMean = TRUE, 
  scale      = 0
)
```

```{r}
tvcm_5_terms <- c(
  "Actual Evapotranspiration",
  "Precipitation", 
  "Min Temperature"
)
```

```{r}
get_tvcoef <- function(plot_vivax, plot_falciparum, labels) {
  nplots_vivax      <- length(plot_vivax)
  nplots_falciparum <- length(plot_falciparum)
  
  assertthat::are_equal(nplots_vivax, nplots_falciparum)
  
  nplots <- nplots_vivax
  
  tvcoef_vivax <-  vector(mode = "list", length = nplots - 1)
  
  for (i in 2:nplots) {
    tvcoef_vivax[[(i-1)]] <- 
      tibble::tibble(
        x        = lubridate::as_datetime(plot_vivax[[i]]$x),
        se       = plot_vivax[[i]]$se,
        fit      = plot_vivax[[i]]$fit,
        term     = labels[(i-1)],
        specie = "vivax"
      )
  }
  
  df_tvcoef_vivax <- dplyr::bind_rows(tvcoef_vivax)
  
  tvcoef_falciparum <-  vector(mode = "list", length = nplots - 1)
  
  for (i in 2:nplots) {
    tvcoef_falciparum[[(i-1)]] <- 
      tibble::tibble(
        x        = lubridate::as_datetime(plot_falciparum[[i]]$x),
        se       = plot_falciparum[[i]]$se,
        fit      = plot_falciparum[[i]]$fit,
        term     = labels[(i-1)],
        specie   = "falciparum"
      )
  }
  
  df_tvcoef_falciparum <- dplyr::bind_rows(tvcoef_falciparum)
  
  df_tvcoef <- dplyr::bind_rows(df_tvcoef_vivax, df_tvcoef_falciparum)
  
  df_tvcoef$term   <- factor(df_tvcoef$term, labels)
  df_tvcoef$specie <- factor(
    x      = df_tvcoef$specie, 
    levels = c("vivax", "falciparum"), 
    labels = c("P. vivax", "P. falciparum")
  )
  
  df_tvcoef
}
```

```{r}
plot_tvcm_by_terms <- function(df_tvcm) {
  
  p <- ggplot(df_tvcm)
  
  # Add grid 
  p <- p + facet_wrap(~ term, ncol = 1, scales = "free", strip = "top")
  
  # Add lines
  p <- p + geom_line(mapping = aes(x = x, y = fit, color = specie), size = 1)
  
  # Add confidence intervals
  p <- p + geom_ribbon(
    mapping = aes(
      x    = x,
      y    = fit,
      ymin = fit + se,
      ymax = fit - se,
      fill = specie
    ),
    alpha = 0.1
  )
  
  # Add PAMAFRO's period layer
  p <- p + geom_rect(
    data    = df_pamafro_period,
    mapping = aes(xmin = start, xmax = end), 
    ymin    = -100, 
    ymax    = 100, 
    alpha   = 0.2
  ) 
    
  # Add PAMAFRO's period lines
  p <- p + geom_vline(
    xintercept = c(df_pamafro_period$start, df_pamafro_period$end), 
    linetype   = "dashed", 
    alpha      = 0.6
  )
  
  # Add y = 0 line
  p <- p + geom_hline(mapping = aes(yintercept = 0), alpha = 0.6)
  
  # Scale date
  p <- p + scale_x_datetime(date_labels = "%Y", date_breaks = "1 year")
  
  # Add labels
  p <- p + labs(y = "b(t)", x = NULL)

  # Format layer
  p <- p + theme(legend.position = "top", legend.title = element_blank())
  
  show(p)
}
```

```{r}
df_tvcm_5<- get_tvcoef(
  plot_vivax_tvcm_5, 
  plot_falciparum_tvcm_5, 
  tvcm_5_terms
)
```

```{r fig.height=9, fig.width=8}
plot_tvcm_by_terms(df_tvcm_5)
```

## Table 2A

```{r}
get_period <- function(x) {
  ifelse(
    x < df_pamafro_period$start,
    "before",
    ifelse(
      x > df_pamafro_period$end,
      "after",
      "during"
    )
  )
}
```

```{r}
get_ci <- function(x) {
  sprintf(
    "%s [%s, %s]", 
    round(x[1], 2), 
    round(x[1] - x[2], 2), 
    round(x[1] + x[2], 2)
  )
}
```

```{r}
table_02_sketch_vivax = htmltools::withTags(
  table(
    class = "display",
    thead(
      tr(
        th(rowspan = 2, "Climate variable"),
        th(colspan = 2, "Before"),
        th(colspan = 2, "During"),
        th(colspan = 2, "After")
      ),
      tr(
        lapply(rep(c("Max", "Min"), 3), th)
      )
    )
  )
)
```

```{r}
df_tvcm_5$period <- apply(df_tvcm_5[, c("x")], 1, get_period)

df_tvcm_5$period <- factor(
  x      = df_tvcm_5$period,
  levels = c("before", "during", "after")
)
```

```{r}
df_tvcm_5 %>% 
  dplyr::group_by(term, period, specie) %>% 
  dplyr::slice_max(fit, n = 1) -> df_tvcm_5_max
```

```{r}
df_tvcm_5 %>% 
  dplyr::group_by(term, period, specie) %>% 
  dplyr::slice_min(fit, n = 1) -> df_tvcm_5_min
```

```{r}
df_tvcm_5_max$extrema <- rep("max", length(df_tvcm_5_max$fit))
df_tvcm_5_min$extrema <- rep("min", length(df_tvcm_5_min$fit))

df_tvcm_5_extrema <- dplyr::bind_rows(df_tvcm_5_max, df_tvcm_5_min)
```

```{r}
df_tvcm_5_extrema$ci <- apply(df_tvcm_5_extrema[, c("fit", "se")], 1, get_ci)
```

```{r}
df_tvcm_5_extrema
```

```{r message=FALSE}
df_tvcm_5_extrema %>% 
  dplyr::filter(specie == "P. vivax") %>% 
  dplyr::select(term, period, extrema, ci) %>%
  tidyr::pivot_wider(
    id_cols     = term,
    names_from  = c(period, extrema),
    values_from = ci,
    names_sep   = "_"
  ) %>% 
  dplyr::select(
    term, 
    before_max, 
    before_min, 
    during_max, 
    during_min, 
    after_max, 
    after_min
  ) -> df_tvcm_5_extrema_vivax
```

```{r}
df_tvcm_5_extrema_vivax %>% 
  tibble::column_to_rownames(var = "term") %>% 
  DT::datatable(
    container  = table_02_sketch_vivax,
    extensions = 'Buttons',
    options    = list(
      dom        = "Bt", 
      scrollX    = TRUE,
      columnDefs = list(list(className = 'dt-center', targets = 6)),
      buttons    = c('copy', 'csv')
    ),
    caption    = "P. vivax"
  )
```

## Table 2B

```{r}
table_02_sketch_falciparum = htmltools::withTags(
  table(
    class = "display",
    thead(
      tr(
        th(rowspan = 2, "Climate variable"),
        th(colspan = 2, "Before"),
        th(colspan = 2, "During"),
        th(colspan = 2, "After")
      ),
      tr(
        lapply(rep(c("Max", "Min"), 3), th)
      )
    )
  )
)
```

```{r message=FALSE}
df_tvcm_5_extrema %>% 
  dplyr::filter(specie == "P. falciparum") %>% 
  dplyr::select(term, period, extrema, ci) %>%
  tidyr::pivot_wider(
    id_cols     = term,
    names_from  = c(period, extrema),
    values_from = ci,
    names_sep   = "_"
  ) %>% 
  dplyr::select(
    term, 
    before_max, 
    before_min, 
    during_max, 
    during_min, 
    after_max, 
    after_min
  ) -> df_tvcm_5_extrema_falciparum
```

```{r}
df_tvcm_5_extrema_falciparum %>% 
  tibble::column_to_rownames(var = "term") %>% 
  DT::datatable(
    container  = table_02_sketch_falciparum,
    extensions = 'Buttons',
    options    = list(
      dom        = "Bt", 
      scrollX    = TRUE,
      columnDefs = list(list(className = 'dt-center', targets = 6)),
      buttons    = c('copy', 'csv')
    ),
    caption    = "P. falciparum"
  )
```

```{r}
df_tvcm_5 %>% 
  dplyr::group_by(period) %>% 
  dplyr::filter(
    x == min(x) | x == max(x)
  )
```
P. vivax

Use max and min

- Actual evapotranspiration

Before PAMAFRO, the actual evapotranspiration effect's had a change of direction from being negative to being positive, going from -0.93 (95% CI: []) at the begining of this period to 0.24 at the end, which means a 74% decrease in the effect's magnitude. 
During PAMAFRO, an opposite change of direction occurred, going from 0.24 to -0.67, which means a 177% increase in the effect's magnitude.

After PAMAFRO, a change in direction in the effect similar to that of the first period happened, going from -0.68 to 0.63, which means a 7% decrease in the effect's magnitude.

- Precipitation

Before PAMAFRO, the precipitation had a negative effect throughout this period, but its magnitude decreased by 98%, going from -1.81 (CI) to -0.04 (CI).

During PAMAFRO, the precipitation continued to have a negative effect on the incidence, but it increased its magnitude 17 times, going from -0.03 (CI) to -0.53 (CI).

After PAMAFRO, a change in direction in the effect similar to that of the first period happened, going from -0.68 to 0.63, which means a 7% decrease in the effect's magnitude.

- Min temperature

## Distributed lag model

### P. vivax

```{r}
tictoc::tic("GAM model fitting")
vivax_tvc_dlm_5 <- mgcv::bam(
  formula = 
    vivax ~ 
    offset(ln_pop2015) + 
    s(district, bs = "re") +
    s(time, by = aet,         bs = "tp", k = 50) +
    s(time, by = aet_lag1,    bs = "tp", k = 50) +
    s(time, by = aet_lag3,    bs = "tp", k = 50) +
    s(time, by = aet_lag6,    bs = "tp", k = 50) +
    s(time, by = aet_lag12,   bs = "tp", k = 50) +
    s(time, by = prcp,        bs = "tp", k = 30) +
    s(time, by = prcp_lag1,   bs = "tp", k = 30) +
    s(time, by = prcp_lag3,   bs = "tp", k = 30) +
    s(time, by = prcp_lag6,   bs = "tp", k = 30) +
    s(time, by = prcp_lag12,  bs = "tp", k = 30) +
    s(time, by = tmin,        bs = "tp", k = 100) +
    s(time, by = tmin_lag1,   bs = "tp", k = 100) +
    s(time, by = tmin_lag3,   bs = "tp", k = 100) +
    s(time, by = tmin_lag6,   bs = "tp", k = 100) +
    s(time, by = tmin_lag12,  bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()

cat("\nConverged?", vivax_tvc_dlm_5$converged, "\n")

mgcv::summary.gam(vivax_tvc_dlm_5)
```

### P. falciparum

```{r}
tictoc::tic("GAM model fitting")
falciparum_tvc_dlm_5 <- mgcv::bam(
  formula = 
    falciparum ~ 
    offset(ln_pop2015) + 
    s(district, bs = "re") +
    s(time, by = aet,         bs = "tp", k = 50) +
    s(time, by = aet_lag1,    bs = "tp", k = 50) +
    s(time, by = aet_lag3,    bs = "tp", k = 50) +
    s(time, by = aet_lag6,    bs = "tp", k = 50) +
    s(time, by = aet_lag12,   bs = "tp", k = 50) +
    s(time, by = prcp,        bs = "tp", k = 30) +
    s(time, by = prcp_lag1,   bs = "tp", k = 30) +
    s(time, by = prcp_lag3,   bs = "tp", k = 30) +
    s(time, by = prcp_lag6,   bs = "tp", k = 30) +
    s(time, by = prcp_lag12,  bs = "tp", k = 30) +
    s(time, by = tmin,        bs = "tp", k = 100) +
    s(time, by = tmin_lag1,   bs = "tp", k = 100) +
    s(time, by = tmin_lag3,   bs = "tp", k = 100) +
    s(time, by = tmin_lag6,   bs = "tp", k = 100) +
    s(time, by = tmin_lag12,  bs = "tp", k = 100),
  family   = nb(),
  data     = train,
  method   = "fREML",
  discrete = TRUE
)
tictoc::toc()


cat("\nConverged?", falciparum_tvc_dlm_5$converged, "\n")

mgcv::summary.gam(falciparum_tvc_dlm_5)
```

## Figure 3

```{r}
tvc_dlm_5_terms <- c(
  "Actual Evapotranspiration (lag = 0)",
  "Actual Evapotranspiration (lag = 1)",
  "Actual Evapotranspiration (lag = 3)",
  "Actual Evapotranspiration (lag = 6)",
  "Actual Evapotranspiration (lag = 12)",
  "Precipitation (lag = 0)",
  "Precipitation (lag = 1)",
  "Precipitation (lag = 3)",
  "Precipitation (lag = 6)",
  "Precipitation (lag = 12)",
  "Min Temperature (lag = 0)",
  "Min Temperature (lag = 1)",
  "Min Temperature (lag = 3)",
  "Min Temperature (lag = 6)",
  "Min Temperature (lag = 12)"
)
```

```{r include=FALSE}
plot_vivax_tvc_dlm_5 <- mgcv::plot.gam(
  x          = vivax_tvc_dlm_5, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0
)

plot_falciparum_tvc_dlm_5 <- mgcv::plot.gam(
  x          = falciparum_tvc_dlm_5, 
  n          = 500, 
  seWithMean = TRUE, 
  scale      = 0
)
```

```{r}
df_tvc_dlm_5 <- get_tvcoef(
  plot_vivax_tvc_dlm_5, 
  plot_falciparum_tvc_dlm_5, 
  tvc_dlm_5_terms
)
```

```{r}
df_tvc_dlm_5
```

```{r}
df_tvc_dlm_5$variable <- factor(
  stringr::str_extract(df_tvc_dlm_5$term, ".*(?=\\s\\()"),
  levels = c("Actual Evapotranspiration", "Precipitation", "Min Temperature")
)

df_tvc_dlm_5$lag <- factor(
  stringr::str_extract(df_tvc_dlm_5$term, "(?<=\\().*(?=\\))"),
  levels = c("lag = 0", "lag = 1", "lag = 3", "lag = 6", "lag = 12")
)
```

```{r fig.height=8, fig.width=12}
df_tvc_dlm_5 %>% 
  ggplot2::ggplot() +
  facet_grid(cols = vars(specie), rows = vars(variable), scales = "free") +
  geom_line(mapping = aes(x = x, y = fit, color = lag), size = 1) +
  geom_ribbon(
    mapping = aes(
      x    = x,
      y    = fit,
      ymin = fit + se,
      ymax = fit - se,
      fill = lag
    ),
    alpha = 0.1
  ) +
  geom_rect(
    data    = df_pamafro_period,
    mapping = aes(xmin = start, xmax = end), 
    ymin    = -60, 
    ymax    = 60, 
    alpha   = 0.2
  ) +
  geom_hline(mapping  = aes(yintercept = 0), alpha = 0.6) +
  geom_vline(
    xintercept = c(df_pamafro_period$start, df_pamafro_period$end),
    linetype   = 'dashed',
    alpha      = 0.8
  ) +
  scale_x_datetime(date_labels = "%Y", date_breaks = "2 year") +
  ggsci::scale_color_npg() +
  ggsci::scale_fill_npg() +
  ylab("b(t)") +
  xlab(NULL) +
  theme(legend.position = "top", legend.title = element_blank())
```

